---
title: "8AR_4C_visualization"
output: html_notebook
---
01/19/2021

lots of preprecessing

# A. first got the matrix.bed files (large on sherlock)
```
[mguo123@sh03-ln07 login /oak/stanford/groups/khavari/users/mguo123/psych_project/data/hicpro_matrix_abs_files]$ wc -l *matrix.bed
   87737453 Astro_B1_5000.matrix.bed
   66167112 Astro_B2_5000.matrix.bed
   65154240 H9-B1_5000.matrix.bed
   54608740 H9-B2_5000.matrix.bed
   56989135 H9D10-B1_5000.matrix.bed
   24774723 H9D10-B2_5000.matrix.bed
   10621516 H9D28-B1_5000.matrix.bed
   11353751 H9D28-B2_5000.matrix.bed
  124558939 H9D2-B1_5000.matrix.bed
  143328326 H9D2-B2_5000.matrix.bed
   66075332 SL-B1_5000.matrix.bed
   75950456 SL-B2_5000.matrix.bed
   77761196 SLC-B1_5000.matrix.bed
  113237474 SLC-B2_5000.matrix.bed
  127742225 SLCD2-B1_5000.matrix.bed
   90346195 SLCD2-B2_5000.matrix.bed
  151907695 SLD2-B1_5000.matrix.bed
  128885574 SLD2-B2_5000.matrix.bed
       1000 test.matrix.bed
 1477201082 total



```

# B. next for each gene symbol we want, run `get_4C.sh` on sherlock
cmd example: `bash get_4C.sh -c 4 -p 113435000 -s NEUROG2` 
make sure to log in setup
transfer files over like sush (run on local computer)
```
cd /Users/mguo123/Documents/pan_omics_psych/data/interim/4C_files
scp mguo123@login.sherlock.stanford.edu:/oak/stanford/groups/khavari/users/mguo123/psych_project/data/hicpro_matrix_abs_files/4C_files/4C_NEUROG* .
```


```
File: get_4C.sh


#!/bin/sh
# get_4c.sh
# enter in gene symbol chrom and position and runs through all samples in this folder to get 4C matrix

cd /oak/stanford/groups/khavari/users/mguo123/psych_project/data/hicpro_matrix_abs_files

samples=(Astro_B1_5000 Astro_B2_5000 H9-B1_5000 H9-B2_5000 H9D2-B1_5000 H9D2-B2_5000 H9D10-B1_5000 H9D10-B2_5000 H9D28-B1_5000 H9D28-B2_5000 SL-B1_5000 SL-B2_5000 SLD2-B1_5000 SLD2-B2_5000 SLC-B1_5000 SLC-B2_5000 SLCD2-B1_5000 SLCD2-B2_5000)


while getopts "c:p:s:" flag
do
    case "${flag}" in
        c) chr=${OPTARG};;
        p) position=${OPTARG};;
        s) sym=${OPTARG};;
    esac
done
echo "chr: $chr";
echo "position: $position";
echo "sym: ${sym}";
#chr=4
#position=113435000
#sym=NEUROG2



for sample in "${samples[@]}"; do
echo $sample

mat_file=$sample.matrix.bed

#chr=4
#position=113435000
#sym=NEUROG2



awk -v position=$position -v chr=$chr  'BEGIN{OFS="\t"} {if ($1==chr && ($2==position|| $5==position)) {if ($2==position){print $1,$2,$3,$4,$5,$6,$7} else {print $4,$5,$6,$1,$2,$3,$7}};}' $mat_file >  4C_files/4C_chr${chr}_${position}_${sym}_${sample}.matrix.bed



done

```

# C. run the following notebook (follows `run_HiChIP_4C_plot.R` from laura D)

## Step 0: setup 

```{r}
setwd("/Users/mguo123/Documents/pan_omics_psych/notebooks")
library(Hmisc)
library(zoo)
library(tidyverse)

# read in dataframes and set save path (FILL IN Accordingly)
save_dir = '../data/processed/4C_plots/'
dir_4C = '../data/interim/4C_files/'
color_df  = read.csv('../data/external/colors.csv', row.names=1)
color_arr = color_df$color_id
names(color_arr) = rownames(color_df)
valid_stat_df = read.table(paste0(dir_4C,'valid_stat.tsv'),sep='\t',
                           row.names=1,header=T)

# set variables (FILL IN PER RUN)
# set variables (FILL IN PER RUN)
samples = c('Astro_B1', 'Astro_B2', 'H9-B1','H9-B2','H9D2-B1','H9D2-B2','H9D10-B1','H9D10-B2',
            'H9D28-B1','H9D28-B2','SL-B1','SL-B2','SLC-B1','SLC-B2','SLD2-B1','SLD2-B2','SLCD2-B1','SLCD2-B2')
gene_sym="CNNM2"
chrom=10
position=104675000 # on file
center_coord=104675000
resolution=5000

# derived variables
### DEBUG if the file reading in doesn't work by 
# list.files(dir_4C,pattern='*bed')
file_prefix = paste0(dir_4C,'4C_chr',chrom, '_', position, '_', gene_sym, '_')
file_suffix = paste0('_',as.character(resolution), '.matrix.bed')
save_file = paste0(save_dir,gene_sym, '_4tissues.pdf')
print(save_file)

```
01/17/2021: bash get_4C.sh -c 4 -p 113435000 -s NEUROG2
01/19/2021: bash get_4C.sh -c 1 -p 212790000 -s ATF3
01/21/2021: bash get_4C.sh -c 5 -p 140180000 -s PCDH1
01/21/2021: bash get_4C.sh -c 8 -p 140180000 -s CHRNA2
  position=27453579 # on file
center_coord=27450000
resolution=5000
01/21/2021: bash get_4C.sh -c 5 -p 140180000 -s PCDH1


## Step 1: Read in and process  the data

process data means:
1. normalization
2. smoothing


```{r}
#
## Define reading table function
# sample : <str> REQUIRED
# norm_method: <str> 'valid', 'sum', default sets table_norm = 1
# rollmean_win: <int> Pick your smoothing window (1 means no smoothing, 2 averages over 2 adjacent bins, etc. )
read_bed = function(sample, norm_method='valid', rollmean_win =2){
  # 1. read bed
  tissue = strsplit(sample, '[_-]')[[1]][1]
  table = read.table(paste0(file_prefix, sample, file_suffix))
  # 2. normalize
  if (norm_method=='valid'){ # from # valid interactions
    table_norm = valid_stat_df[sample, "X"]
  }
  else if (norm_method=='sum'){ # sum/1000
    table_norm = sum(table[,7])/1e3
  }
  else{ # no norm
    table_norm = 1
  }
  
  # Process table
  counts_left=table[table[,2]== center_coord,c(5,7)]
  counts_right=table[table[,5]== center_coord,c(2,7)]
  colnames(counts_left) = c(1,2)
  colnames(counts_right) = c(1,2)
  counts_merge=rbind(counts_left, counts_right)
  counts_merge_ord = counts_merge[order(counts_merge[,1]),]
  counts_merge_ord[,2] = counts_merge_ord[,2]/table_norm
  counts_merge_ord[,1] = counts_merge_ord[,1] + resolution/2
  line = rollmean(counts_merge_ord[,2], rollmean_win,fill=c(0,0,0))
  
  
  df = data.frame(x = counts_merge_ord[,1], y= line, sample = sample, tissue = tissue)
  
  return (df)
}

# read in chosen samples
tables = lapply(samples, function(x) read_bed(x, rollmean_win=3))
table_all = bind_rows(tables)
table_all_avg = table_all%>%
  group_by(x,tissue)%>%
  summarise(y=mean(y))

```



## Step 2:  Plot

```{r}
# Plot
# plotting variables
xlim_minus=4e5 # bp upstream of anchor
xlim_plus=4e5 # bp downstream of anchor
xlim_par= c((center_coord)-xlim_minus,(center_coord)+xlim_plus)
ylim_par=c(-0.01,0.2) # y axis limits
lwd_par = .5 # line thickness 
xlab="Position on chr"
ylab="Filtered reads"
plot_name=paste("Virtual 4C anchor 5kb bin for gene: ",gene_sym, ' at chr',chrom, ':',center_coord,sep="")


# actual plot 
ggplot(table_all_avg, aes(x=x, y=y, color=tissue))+geom_line(lwd=lwd_par)+  
  scale_colour_manual(values=color_arr)+
  labs(title=plot_name,
        x =xlab, y = ylab)+
  xlim(xlim_par) + ylim(ylim_par)+
  theme_classic()+ theme(plot.title = element_text(size=10))+
  geom_vline(xintercept = center_coord + resolution/2,color='gray',lty=2)+
  geom_vline(xintercept = 104426177,color='red',lty=2)+
  geom_vline(xintercept = 104767185,color='red',lty=2)+
  geom_vline(xintercept = 104837816,color='red',lty=2)+
  geom_vline(xintercept = 104941112,color='red',lty=2)+
  geom_vline(xintercept = 104953007,color='red',lty=2)

# saving plot
ggsave(save_file) # CAN CHANGE width and height
save_file
```

